---
title: "Databases lab"
author: "Jeff Leek"
date: "July 12, 2016"
output: html_document
---

1. Download and load the nycflights data with the command `install.packages('nycflights13')` and `library(nycflights13)`.

```{r}
install.packages("nycflights13")
pacman::p_load(nycflights13)
```


2. Use the `pryr` package to figure out the size of the `flights` object.

```{r}
object_size(flights)
```


3. Create a sqlite database, then add a table "flights" with the flights data from this package.

```{r}
my_db <- dplyr::src_sqlite("my_db.sqlite3", 
                    create = TRUE)
flights_sqlite <- dplyr::copy_to(my_db, 
                        flights, temporary = FALSE)
dplyr::src_tbls(my_db)
tbl(my_db,"flights")
```


4. Inspect the tables using the `src_tbls` command to make sure the copying happened correctly.

```{r}
src_tbls(my_db)
```


5. Find the average delay time for American Airlines (hint: the abbreviation is AA).

```{r}
test<- my_db %>% 
  tbl("flights")%>%
  filter(carrier == "AA") %>% 
  collect() %>% 
  summarise(avg.delay = mean(arr_delay, na.rm = T))
```


6. Can you add a variable for delay time to the database?

```{r}
start <- Sys.time()
new.data <- 
  my_db %>% 
  tbl("flights")%>%
  group_by(carrier) %>% 
  summarise(total.delay = (dep_delay + arr_delay)) %>% 
  collect()



Sys.time() - start
```


7. How long does it take to collect the results of your computation for 5? 

8. Can you figure out the average delay time for each airline? 

```{r}
my_db %>% 
  tbl("flights")%>%
  group_by(carrier) %>% 
  summarise(avg.delay = mean(arr_delay, na.rm = T)) %>% 
  collect()

```

