---
title: "Merging Lab"
author: "Data Wrangling in R"
output: html_document
---

1. Download the samples data on the 1,000 genomes project from our website (https://sisbid.github.io/Module1/labs/1000genomes.xlsx).

```{r}
samples <- read_excel("labs/1000genomes.xlsx", sheet = 1)
sequence <- read_excel("labs/1000genomes.xlsx", sheet = 4, skip = 1)
sequence <- sequence[, 1:7]

```

2. Open the sheet and look at the information listed. Go to the sheet *Final Phase Sequence Data*

3. Read in the first and fourth sheets (you may want to use the `readxl` package). Keep only the low coverage data from the 4th sheet. 

4. Look at the cheetsheet on merging:  http://stat545.com/bit001_dplyr-cheatsheet.html


5. Try merging the data sets with `inner_join`, `semi_join`, `left_join`,`full_join`. 
```{r}
inner <- inner_join(samples, sequence)
semi <- semi_join(samples, sequence)
left <- left_join(samples, sequence)
full <- full_join(samples, sequence)
```


6. What are the dimensions of the resulting data sets? 

7. The full join and and left join have the same dimensions. Are they exactly identical? 

```{r}
identical(full, inner)
```


8. Reverse the order of that sets on `left_join`. Do you get the same result? 

```{r}
left.rev <- left_join(sequence, samples)
```


9. Try `anti_join`. What happens? Read here for more on filtering joins: http://r4ds.had.co.nz/relational-data.html

```{r}
anti <- anti_join(samples, sequence)
itna <- anti_join(sequence, samples)
```

