---
title: "Web + API Lab"
author: ""
date: "July 6, 2015"
output: html_document
---

### Web scraping

In this exercise we will be scraping data from the 2019 World Cup (that was totally just won by the amazing US Women's team :)). 



1. Go to the website https://www.espn.com/soccer/league/_/name/fifa.wwc

```{r}
women.url <- "https://www.espn.com/soccer/league/_/name/fifa.wwc"
htmlfile <- read_html(women.url)
```

2. View the source to see how the page is structured, can you find the standings table in the HTML code?
3. Inspect the element and try to find the Xpath for the table for the standings. 
4. Adapt the code we learned to scrape the standings table. 

```{r}

nds <- html_nodes(htmlfile,                
                 xpath='//*[@id="main-container"]/div/section[2]/article[2]/div/table')

dat <- html_table(nds)
dat <- as.data.frame(dat)
head(dat)
```



6. Go to https://www.bioconductor.org/packages/release/BiocViews.html#___Software
7. Try to find the table of packages in the source code. 
```{r}

htmlfile <- read_html("https://www.bioconductor.org/packages/release/BiocViews.html#___Software")
nds <- html_nodes(htmlfile,                
                 xpath='//*[@id="biocViews_package_table"]')

dat <- html_table(nds)
dat <- as.data.frame(dat)
head(dat)
```



8. Can you figure out why it is or isn't there? 


### APIs

Let's play around with the Ensembl API (http://rest.ensembl.org/) for extracting informtion. This is an open API so you don't need a developer account. 

1. Download information on _Oct4_ from Ensembl using this call:
```{r,eval=FALSE}
library(jsonlite)
oct4_dat = fromJSON("http://rest.ensembl.org//lookup/id/ENSG00000204531?content-type=application/json;expand=1")
```

2. Explore the `oct4_dat` variable that results. Look particularly at `oct4_dat$Transcript` column by column with the commands `oct4_dat$Transcript`

4. The Ensembl ID for the gene _P53_ is ENSG00000141510. How do you think you'd change the command above to look up this gene?

```{r}
p53_dat = fromJSON("http://rest.ensembl.org//lookup/id/ENSG00000141510?content-type=application/json;expand=1")
```


5. Based on the Ensembl docs how would you look up the ensembl id for another gene like _BRAF_? 

To look up the BRAF gene information we would. 

1. Look at the docs under GET species lookup http://rest.ensembl.org/documentation/info/symbol_lookup. 
2. We'd look at the Example Request: `/lookup/symbol/homo_sapiens/BRCA2?content-type=application/json;expand=1` and modify it with our gene like so: 
`/lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1`
3. We'd crete the url using the base url like this: `http://rest.ensembl.org//lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1`
4. We'd request the information with `fromJSON`
```{r}
braf_info = fromJSON("http://rest.ensembl.org//lookup/symbol/homo_sapiens/BRAF?content-type=application/json;expand=1")
braf_info$id
```

